Community-Conditioned Language Models

### 2021-05-04 paper/paper.tex
notes

\newcommand\Ppl{\mathsf{Ppl}}
\newcommand\IG{\mathsf{IG}}
\newcommand\Ind{\mathsf{Ind}}
\DeclareMathOperator*{\avg}{average}
\newcommand\BibTeX{B\textsc{ib}\TeX}

@@

TODO: outline of the paper.

@@

We experiment with two kinds of model architecture: a simple
unidirectional LSTM \citep{Hochreiter1997} and a Transformer
\citep{Vaswani2017}.  In either case, the model is organized as a
standard $n$-layer neural sequence encoder (we set $n=3$ for all
models). As conventional, word tokens are first converted to vectors
using a trainable embedding layer. At the other end, word tokens are
predicted using a softmax projection layer. The set of word tokens is
a pre-determined vocabulary.  These encoders form the core of the
of our models, as such do not take community into account, and we call them
unconditioned models.

To condition the model (and obtain community-conditioned language
models --- CCLM), we add a \emph{community embedding} parameter to the
models.  This parameter is concatenated with the hidden layer of the
sequence encoder, for some layer $l_c \leq n$, and passed through a
linear layer which projects the resulting vector back to the original
hidden layer size.  For $l_c = n$, the output of this linear layer is
passed directly to the softmax function, just as the final hidden
layer of the sequence encoder is in other models.  For $l_c=0$, the
community embedding is concatenated with the token embedding.  For
this reason, we set the hidden size of the sequence encoder and the
size of the token to be embedding equal for all models.

@@

We investigate linguistic variation across various communities present
on the social media website Reddit.\footnote{Comments were obtained
  from the archive at \url{https://pushshift.io/}.
  \cite{Baumgartner2020}.}
%
Reddit is divided into forums called \textit{subreddits}, 
which are typically organized around a topic of interest. 
Users create \textit{posts}, which consist of a link, image, 
or text, along with a \emph{comment} section. 
Comments are threaded: a comment can be made directly on a post,
or appear as a reply to another comment.
%
Hereafter we refer to such comments as ``messages'', to match our
convention in mathematical formulas: the letter $c$ stands for a
community, and $m$ stands for a message.

@@

Messages were preprocessed as follows: 
We excluded the content of block quotes, code blocks, and tables.
We removed any markup (formatting) commands from the remaining text, extracting only rendered text.
We tokenized the messages using the default English model for the SpaCy tokenizer Version 2.2.3 \citep{Honnibal2017}
and lower-cased all tokens.

@@

We combined the 46 community corpora and split the combined corpus,
reserving 10\% for testing, 10\% for validation and the rest for training.
Splits were stratified by community so that each community had \num{5000} examples
for testing, \num{5000} for evaluation, and \num{40000} for training.

In the remainder of the paper we use $M$ to refer to messages in the
test set, and $M_j$ for the partition of the test set originating from
community $c_j$.

We use a vocabulary size of \num{40000} tokens, including a special
out-of-vocabulary token.  The vocabulary consists of the most frequent
tokens at the message level (across all communities), meaning that we
only count each token once per message.  The main reason for this
choice is to avoid over-counting otherwise rare tokens that were
repeated many times in a single message.

@@

We trained the models until the validation loss stopped decreasing for
two epochs in a row, and used the weights from the epoch with the
smallest validation loss for testing.  In general, transformer models
trained for about half as long as the LSTM models (test epoch,
\cref{tab:model-results}).

@@

\subsubsection{Losses of CCLM}

We measure the perplexity of each message in the test set. 
For this, we use the negative exponential of the mean cross-entropy loss 
of the language model for a given message and average over a set of messages, $M$.
\[\Ppl_M = e^{\avg_{m\in M} H(m)}\]
As one would expect, 
the conditioned models have lower perplexity
than their respective unconditioned baseline models (\cref{tab:model-results}),
but this effect is different in magnitude
for models with different community embedding depths.
The best transformer model ($l_c=3$)
incorporates the community information last,
after all the sequence encoder layers.
For the LSTM encoder, 
the model that concatenates the community embedding after the first encoder layer ($l_c=1$),
performs almost as well as the model with $l_c=3$, but that model is still the best.
There is much less variation in performance across different values of $l_c$ for the LSTM.

@@

We also consider model performance stratified by community; that is,
$\Ppl_{M_j}$, where $M_j$ is the test-set messages from community
$c_j$.  We observe a lot of variation in baseline perplexity
across communities, with $\Ppl_{M_j}$ ranging from \num{34.47} to
\num{66.43} for the LSTM model (see table 1 in the supplementary materials).

\subsection{Information gain of the CCLM}

Next, we consider the \emph{information gain} of each CCLM over its baseline
un-conditioned counterpart with the same sequence encoder architecture.
For a given message, information gain is defined as the difference
between the cross-entropy of the unconditioned model and the conditioned model:
\[H_{LM}(m) - H_{CCLM}(m)\]
For a set of messages, $M$, we consider the average information gain
in exponential space (as a ratio of perplexities):
\[\IG_M = \frac{e^{\avg_M(H_{LM}(m))}}{e^{\avg_M(H_{CCLM}(m))}}\]
%\jp{What we had before: average entropy gain is $H(m)-H_0(m)$, and
  %then taking the average over all messages:
  %$avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m)) = H - H₀$. Taking the exponential
  %gives a ratio of perplexity.  Here it seems that you have commuted
  %exponential and average. It is rather weird to do average in the
  %exponential space without normalising to a probability.

  %Compare: $e^{avg(H(m))} / e^{avg(H_0(m))}$ with $avg \frac{e^{H(m)}}{e^{H₀(m)}}$.
%}

While the absolute performance of the LSTM models is better,
we find that the best transformer CCLM has higher information gain 
than the best LSTM (\cref{tab:model-results}).

The conditioned models also perform differently across different communities---%
for some communities, $\IG_{M_j}$ around \num{1}, or even slightly lower,
meaning that the CCLM has no advantage over the unconditioned baseline for that community.
For other communities it is higher, meaning that the CCLM performs better than 
the baseline for those communities (\cref{fig:comm-info-gain}).

\begin{figure}
\begin{tikzpicture}
  \pplbycommbox
\end{tikzpicture}
\caption{
  Distribution of $\Ppl_{M_j}$ for each model architecture.
}
\label{fig:comm-info-gain}
\end{figure}

This variation in information gain across communities is not 
related to perplexity, however. 
Some of the most linguistically complex communities benefit a lot
from community conditioning, while others do not benefit at all. 
Similarly, among the less linguistically complex communities 
there some have high information gain, and some have information
gain of \num{1} or less (\cref{fig:ppl-info-gain}).

\begin{figure}
\begin{tikzpicture}
\begin{axis}
\commscatter[]{lstm-3-1_lm_ppl}{lstm-3-1_info_gain}
\end{axis}
\end{tikzpicture}
\caption{
  Average community perplexity (x-axis) versus information gain (y-axis)
  for the LSTM model with $l_c = 1$ (colors based on \cref{fig:comm-pca}).
}
\label{fig:ppl-info-gain}
\end{figure}

@@

\subsection{Latent language model-based community classifiers}

@@

We can characterize the quality the LMCC with a confusion matrix
$C$, defined by $C_{ij} = P(c=c_i \mid m)$ for a given (test set) message $m$ 
from community $c_j$.
\[C_{ij} = \avg_{m\in M_j}P(c=c_i \mid m)\]
We observe that the diagonal dominates the matrix: the LMCC generally finds the right community for a message (\cref{fig:confusion}).
%
\begin{figure}
\begin{tikzpicture}
  \confusionsingleplot{lstm-3-1}
\end{tikzpicture}
\caption{LMCC confusion matrices for the LSTM model with $c=1$.}
\label{fig:confusion}
\end{figure}
%
But we want a more quantitative measure, corresponding to the ability of the LMCC to correctly
recognize messages coming from a given community $c_j$. We could use
something like an F-score, but we prefer to use a perplexity-like metric,
which is has stronger information-theoretic grounds.
To this end, we define the \emph{linguistic indiscernibility} of community $c_j$, 
noted $\Ind_j$. 
First, note that each row of $C$ is a probability distribution,
since it is computed as the average of $P(c=c_i \mid m)$ for $m\in M_j$.
The exponential of the entropy of the $j$th row of \(C\) is therefore a perplexity score,
whose maximum is the total number of communities (46 for us).
Thus we define $\Ind_j$ by scaling this perplexity by the number of communities:
\[\Ind_j = \frac{e^{-\sum_i C_{ij} log(C_{ij})}}{|C_j|}\]
This way $Ind_j=0$ means that the LMCC is completely certain of its
choice, while $Ind_j=1$ means that equal probability is given to all
predictions.  In general $\Ind_j$ is a score of uncertainty of the
LMCC, given messages from a community $c_j$.

Indiscernibility of a given community is a robust measure: it is
almost perfectly correlated across CCLMs (such as transformers, LSTM
and various values of $l_c$).  In fact, we observe a correlation
between the indiscernibility measures by every pair of models with
Pearson's $r \geq 0.98$ and $p < 0.0001$ (see, for example,
\cref{fig:indisc-corr}).
%
\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
    \commscatter{lstm-3-1_indisc}{transformer-3-3_indisc}
    \end{axis}
  \end{tikzpicture}
  \caption{%
    Correlation between indiscernibility for 
    LSTM ($c = 1$) and transformer ($c = 3$) models
    (Pearson's $r = 0.99$, $p < 0.0001$).
  }
  \label{fig:indisc-corr}
\end{figure}
%
However, we observe that community-wise perplexity is only weakly (if at all) 
correlated with community indiscernibility under the corresponding latent 
LMCC (see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm-lmcc-ppl}).
In other words, the linguistic complexity of a given community
is a poor predictor of how distinctive its language is.
%
\begin{table}
  \small
  \centering
  \input{floats/ind_corrs.tex}
  \caption{ Pearson's $r$ correlation coefficient between community
    indiscernibility ($\Ind_j$) and two different predictors: CCLM
    perplexity on messages from community $c_j$ ($\Ppl_{M_j}$), and
    CCLM information gain on messages from $c_j$ ($\IG_{M_j}$).  }
  \label{tab:ind-corrs}
\end{table}
%
\begin{figure*}
  \begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1},
    legend style={anchor=north, legend columns=1, font=\tiny},
    every node near coord/.append style={font=\sffamily\tiny}%
  ]
    \nextgroupplot
    \commscatter[]{lstm-3-1_lm_ppl}{lstm-3-1_indisc}
    \nextgroupplot
    \commscatter[]{lstm-3-1_info_gain}{lstm-3-1_indisc}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
  \end{tikzpicture}
  \caption{%
    Average community indiscernibility (y-axis), 
    in relationship to linguistic complexity (left) and 
    CCLM information gain (right). Shown here for the LSTM
    model with $l_c=1$. See \cref{tab:ind-corrs} for 
    Pearson's r correlations for all models.
  }
  \label{fig:cclm-lmcc-ppl}
\end{figure*}

@@

Computing this distance is known as the orthogonal Procrustes problem~\citep{Gower2004}. Its
solution is
\[d(L,S) = \mathsf {Tr} Σ\] where the matrix \(Σ\) is obtained by the singular value
decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
the directions of correlation respectively of \(L\) and \(S\). The
singular values \(\sigma_i\) (diagonal elements of \(Σ\)), say how much
correlation there is between the directions \(U_i\) and \(V_i\).

Further, we can arrange the solution so that \(σ_i < σ_j\) iff \(i < j\). That
is, the largest singular value \(σ_0\) corresponds to the principal
directions of correlation (\(U_0\), \(V_0\)), \(σ_1\) to the second
principal direction, etc.

The \(d(L,S)\) metric varies between \(0\) (perfect independence) and
\(n\) (perfect correlation, obtained for example if \(L=S\)), where
$n$ is the number of community embeddings considered.

To get a more precise sense of what value of \(d(L,S)\) indicates significant correlation,
We have computed experimentally \(d(L,S)\) for 
\(L\) and \(S\) chosen at random and 16 dimensions. We find
a mean of TODO and a standard deviation of TODO.

Thus \(d(L,S)\) below \(μ-3σ\) can be safely considered to indicate a
statistically significant correlation between \(L\) and
\(S\). However, below, we will also give a sense of \emph{how} the
correlation is manifested by analysis of the two principal components
of correlation in the linguistic embeddings, \(U_0\) and \(U_1\).

@@

\begin{figure*}[t]
\begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1},
    yticklabels={,,}, xticklabels={,,},
    legend style={anchor=north, legend columns=1, font=\sffamily\tiny},
  ]
    \nextgroupplot
    \commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
    \nextgroupplot
    \commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
\end{tikzpicture}
\caption{Plot of embeddings along the 2 principal dimensions of correlation between .... TODO of an LSTM (left, $c=1$) and transformer (right, $c=3$) model.
Color correspond the 5-means clustering of .... TODO}
\label{fig:comm-pca}
\end{figure*}

>>>>

\newcommand\Ppl{\mathsf{Ppl}}
\newcommand\IG{\mathsf{IG}}
\newcommand\Ind{\mathsf{Ind}}
\newcommand\LM{\mathrm{LM}}
\newcommand\CCLM{\mathrm{CCLM}}
\DeclareMathOperator*{\avg}{average}
\newcommand\BibTeX{B\textsc{ib}\TeX}

@@

%TODO: outline of the paper.

@@

We experiment with two kinds of model architecture: a simple
unidirectional LSTM \citep{Hochreiter1997} and a Transformer
\citep{Vaswani2017}.  In either case, the model is organized as a
standard $n$-layer neural sequence encoder (we set $n=3$ for all
models). As is convention, a sequence of word tokens is first converted to 
vectors using a trainable embedding layer over a pre-deterimend vocabulary.
At the other end, word tokens are predicted using a softmax projection 
layer. What we have described so far does not take community into account
and as such we call them \emph{uncodnitioned models}, but the same
encoder architecture also forms the core of our conditioned models.

To condition the model , we add a \emph{community embedding} parameter.
This parameter is concatenated with the hidden layer of the
sequence encoder, at some layer $l_c \leq n$, and passed through a
linear layer which projects the resulting vector back to the original
hidden layer size.  For $l_c = n$, the output of this linear layer is
passed directly to the softmax function, just as the final hidden
layer of the sequence encoder is in other models.  For $l_c=0$, the
community embedding is concatenated with the token embedding.  For
this reason, we set the hidden size of the sequence encoder and the
size of the token embedding to be equal for all models.

%TODO: Diagram of model?

@@

We investigate linguistic variation across various communities 
from the social media website Reddit.\footnote{Comments were obtained
  from the archive at \url{https://pushshift.io/}.
  \cite{Baumgartner2020}.}
%
Reddit is divided into forums called \textit{subreddits}, 
which are typically organized around a topic of interest. 
Users create \textit{posts}, which consist of a link, image, 
or text, along with a \emph{comment} section. 
Comments are threaded: a comment can be made directly on a post,
or appear as a reply to another comment.
%
Hereafter we refer to such comments as ``messages'', to match our
convention in mathematical formulas: the letter $c$ stands for a
community, and $m$ stands for a message.

@@

Messages were preprocessed as follows: 
We excluded the content of block quotes, code blocks, and tables.
We removed any markup (formatting) commands from the remaining text, 
extracting only rendered text.
We tokenized the messages using the default English model for the SpaCy tokenizer 
Version 2.2.3 \citep{Honnibal2017}.

@@

We use a vocabulary size of \num{40000} tokens, including a special
out-of-vocabulary token.  The vocabulary consists of the most frequent
tokens across all communities.

@@

We trained the models until the validation loss stopped decreasing for
two epochs in a row, and used the weights from the epoch with the
smallest validation loss for testing.  In general, transformer models
trained for about half as long as the LSTM models (\cref{tab:model-results}, 
test epoch).

@@

In this section, we report the performance of the conditioned and un-conditioned
models on the test set.
First, we define two performance metrics, perplexity and information gain.
In the following, we use $M$ to refer to messages in the
combined test set, and $M_j$ for the partition of the test set originating from 
community $c_j$.

\paragraph{Perplexity}

For a given model let $H(m)$ be the model's cross-entropy loss,
averaged over tokens in $m$.
We define the perplexity on a set of messages, $M$,
to be the exponential of the model's average cross-entropy loss:
\[\Ppl_M = e^{\avg_{m\in M} H(m)}\]

\paragraph{CCLM Information Gain}

We also consider the information gain of each CCLM over its baseline
un-conditioned counterpart with the same sequence encoder architecture.
For a given message, information gain is defined as the difference
between the cross-entropy of the unconditioned model and the conditioned model:
\[H_{\mathrm{LM}}(m) - H_{\mathrm{CCLM}}(m)\]
For a set of messages, $M$, we consider the average information gain
in exponential space (as a ratio of perplexities):
\[\IG_M = \frac{e^{\avg_M(H_{\mathrm{LM}}(m))}}{e^{\avg_M(H_{\mathrm{CCLM}}(m))}}\]
%\jp{What we had before: average entropy gain is $H(m)-H_0(m)$, and
  %then taking the average over all messages:
  %$avg(H(m)-H_0(m)) = avg(H(m))-avg(H_0(m)) = H - H₀$. Taking the exponential
  %gives a ratio of perplexity.  Here it seems that you have commuted
  %exponential and average. It is rather weird to do average in the
  %exponential space without normalising to a probability.
  %Compare: $e^{avg(H(m))} / e^{avg(H_0(m))}$ with $avg \frac{e^{H(m)}}{e^{H₀(m)}}$.
%}
\bn{I think this is stated correctly now (not the average in exp space)}

\paragraph{}

Unsurprisingly, the conditioned models mostly have lower perplexity 
than their respective unconditioned baseline models, 
meaning information gain above one (\cref{tab:model-results}).
Overall, the LSTM models perform better than the transformers, 
although the best transformer has higher information gain than the best LSTM.

The effect of $l_c$, the depth of the community embedding,
is also different across architectures.
For the LSTM encoder, 
the best model concatenates the community embedding after the first encoder layer ($l_c=1$),
but all of the conditioned models perform similarly well.
For the transformer, the best model incorporates the community information
first, concatenating it directly to the word vectors ($l_c=0$).
It performs similarly to the model that only integrates the community information
after all all the transformer layers ($l_c=3$),
but the two middle-layer models actually perform worse than the unconditioned model
(with $\IG_M < 1$).

@@

We also consider performance stratified by community; that is,
$\Ppl_{M_j}$ and $\IG_{M_j}$, 
where $M_j$ is the set of messages originating from community $c_j$ 
(\cref{fig:comm-stratified-box}).
We observe a lot of variation in baseline perplexity
across communities, with $\Ppl_{M_j}$ ranging from \num{3.67} to
\num{93.58} for the best LSTM model . %TODO make long table
While the absolute performance ($\Ppl_{M_j}$) of the LSTM models is better,
we jhat the best transformer models have somewhat higher information gain
than their LSTM counterparts.

\newcommand{\modelboxplot}[3]{
  \addplot+[
    boxplot={draw position=#3, box extend=0.3}, 
    draw=#2, mark=*, mark options={fill=#2, scale=0.5}, solid, fill=#2!10,
    area legend] 
    table [y=#1] {floats/comm.csv};
}
\begin{figure}[t]
\begin{tikzpicture}
  \begin{axis}[name=ig,
    height=7cm,width=8cm,
    boxplot/draw direction=x, axis x line*=bottom, axis y line*=left,y axis line style={opacity=0},
    %xmax=1.6,
    ylabel=$l_c$, xlabel=$\IG_{M_j}$,
    legend style={font=\tiny,at={(1,1)},anchor=north east}, legend columns=2
    ]
    \addplot[thin, red, samples=2] coordinates {(1,-0.5)(1,3.5)};
    \foreach \lc in {0, ..., 3}{
      \modelboxplot{lstm-3-\lc_ig}{violet}{\lc+0.2}
      \modelboxplot{transformer-3-\lc_ig}{teal}{\lc-0.2}
    \legend{,LSTM, Transformer}
    }
  \end{axis}
  \begin{axis}[name=ppl,
    anchor=north west,at={($(ig.south west)-(0,1.1cm)$)},
    width=8cm, height=2.5cm, 
    boxplot/draw direction=x, axis x line*=bottom, hide y axis,
    xlabel=$\Ppl_{M_j}$,
    ]
    \modelboxplot{lstm-3_ppl}{violet}{0.2}
    \modelboxplot{transformer-3_ppl}{teal}{-.2}
  \end{axis}
\end{tikzpicture}
\caption{
   Average model performance by community. 
   Also see table 1 in the supplementary materials
   for a tabular version of these results with community names.
}
\label{fig:comm-stratified-box}
\end{figure}

The conditioned models also perform differently across different communities---%
even among the best models, some communities have $\IG_{M_j} < 1$,
meaning that the CCLM performs worse than the unconditioned baseline for 
messages from that community.
For other communities $\IG_{M_j}$ is much higher, meaning that the CCLM performs better  
(\cref{fig:comm-info-gain}).\footnote{
  Some of the communities with consistently high $IG_{M_j}$ 
  across all models are primarily non-English, 
  but surprisingly, not the three most extreme
  outliers. There are /r/counting, /r/friendsafari, and /r/Fireteams, 
  the later two of which are places where people coordinate to play video games
  together. The coordination has highly strucutred and regular formats, which 
  are presumably conventional to these communities.}

The variation in information gain across communities is not 
related to perplexity, however (\cref{fig:ppl-info-gain}).

\begin{figure}
\begin{tikzpicture}
\begin{axis}[%
    axis x line*=bottom, axis y line*=left,
    xlabel=$\Ppl_{M_j}$, ylabel=$\IG_{M_j}$,
    width=7cm,
  ]
  \addplot+[%
    only marks,
    mark=*,
    mark options={scale=0.35, solid, color=violet!80, fill=violet},
    font=\tiny,
  ] table [x={lstm-3_ppl}, y={lstm-3-3_ig}] {floats/comm.csv};
\end{axis}
\end{tikzpicture}
\caption{
  Average community perplexity (x-axis) versus information gain (y-axis)
  for the best LSTM model ($l_c = 1$).
}
\label{fig:ppl-info-gain}
\end{figure}

@@

\section{Latent language model-based community classifiers}

@@

We can characterize the quality the LMCC with a confusion matrix
$C$, defined by $C_{ij} = P(c=c_i \mid m)$ for a given (test set) message $m$ 
from community $c_j$.
\[C_{ij} = \avg_{m\in M_j}P(c=c_i \mid m)\]
We observe that the diagonal dominates the matrix: the LMCC generally finds the right community for a message (\cref{fig:confusion}).

\begin{figure}
  \includegraphics[width=\linewidth]{floats/confusion_lstm-3-1}
\caption{
  Log-probability LMCC confusion matrices for the LSTM model with $l_c=1$. 
  Communities are sorted by indiscernability.}
\label{fig:confusion}
\end{figure}

But we want a more quantitative measure, corresponding to the ability of the LMCC to correctly
recognize messages coming from a given community $c_j$. We could use
something like an F-score, but we prefer to use a perplexity-like metric,
which is has stronger information-theoretic grounds.
To this end, we define the \emph{linguistic indiscernibility} of community $c_j$, 
noted $\Ind_j$. 
First, note that each row of $C$ is a probability distribution,
since it is computed as the average of $P(c=c_i \mid m)$ for $m\in M_j$.
The exponential of the entropy of the $j$th row of \(C\) is therefore a perplexity score,
whose maximum is the total number of communities (46 for us).
Thus we define $\Ind_j$ by scaling this perplexity by the number of communities:
\[\Ind_j = \frac{e^{-\sum_i C_{ij} log(C_{ij})}}{|C_j|}\]
This way $Ind_j=0$ means that the LMCC is completely certain of its
choice, while $Ind_j=1$ means that equal probability is given to all
predictions.  In general $\Ind_j$ is a score of uncertainty of the
LMCC, given messages from a community $c_j$.

Indiscernibility of a given community is a robust measure: it is
almost perfectly correlated across CCLMs (such as transformers, LSTM
and various values of $l_c$).  In fact, we observe a correlation
between the indiscernibility measures by every pair of models with
Pearson's $r \geq 0.96$ and $p < 0.0001$.

However, we observe that community-wise perplexity is only weakly (if at all) 
correlated with community indiscernibility under the corresponding latent 
LMCC (see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm-lmcc-ppl}).
In other words, the linguistic complexity of a given community
is a poor predictor of how distinctive its language is.
%
\begin{table}
  \small
  \centering
  \input{floats/ind_corrs.tex}
  \caption{ Pearson's $r$ correlation coefficient between community
    indiscernibility ($\Ind_j$) and two different predictors: CCLM
    perplexity on messages from community $c_j$ ($\Ppl_{M_j}$), and
    CCLM information gain on messages from $c_j$ ($\IG_{M_j}$).  }
  \label{tab:ind-corrs}
\end{table}

\begin{figure*}
  \begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1},
    legend style={anchor=north, legend columns=1, font=\tiny},
    every node near coord/.append style={font=\sffamily\tiny},%
    axis x line*=bottom, axis y line*=left,
  ]
  \nextgroupplot[xlabel=$\Ppl_{M_j}$, ylabel=$\Ind_j$]
      \addplot+[%
        only marks,
        mark=*,
        mark options={scale=0.35, solid, color=violet!80, fill=violet},
        font=\tiny,
      ] table [x={lstm-3-1_ppl}, y={lstm-3-1_indisc}] {floats/comm.csv};
  \nextgroupplot[xlabel=$\IG_{M_j}$, ylabel=$\Ind_j$]
      \addplot+[%
        only marks,
        mark=*,
        mark options={scale=0.35, solid, color=violet!80, fill=violet},
        font=\tiny,
      ] table [x={lstm-3-1_ig}, y={lstm-3-1_indisc}] {floats/comm.csv};
  \end{groupplot}
  \end{tikzpicture}
  \caption{%
    Community indiscernibility (y-axis), 
    in relationship to linguistic complexity (left) and 
    CCLM information gain (right). Shown here for the LSTM
    model with $l_c=1$. See \cref{tab:ind-corrs} for 
    Pearson's r correlations for all models.
  }
  \label{fig:cclm-lmcc-ppl}
\end{figure*}

@@

Computing \(d(L,S)\) is known as the orthogonal Procrustes problem~\citep{Gower2004}. The
solution is
\[d(L,S) = \mathsf {Tr} Σ\] where the matrix \(Σ\) is obtained by the singular value
decomposition \(U^TΣV = LS^T\). Additionally the vectors of \(U\) and \(V\) give
the directions of correlation respectively of \(L\) and \(S\). Each
singular \(\sigma_i\) (elements of the diagonal matrix \(Σ\)), give a measure of how much
correlation there is between the directions \(U_i\) and \(V_i\).

Further, we can apply a permutation to \(U\),\(V\) and \(Σ\) such that \(σ_i < σ_j\) iff \(i < j\). That
is, the largest singular value \(σ_0\) then corresponds to the principal
directions of correlation (\(U_0\), \(V_0\)), \(σ_1\) to the second
principal direction, etc.

The \(d(L,S)\) metric range from \(0\) (corresponding to perfect
independence) to \(n\) (corresponding to perfect correlation, obtained
for example if \(L=S\)), where $n=510$ is the number of communities
considered.

To get a more precise sense of what value of \(d(L,S)\) indicates significant correlation,
We have computed experimentally \(d(L,S)\) for 
\(L\) and \(S\) chosen as random 16 dimensions. We find
a mean of TODO and a standard deviation of TODO.

TODO: !!! we should take a 16 × 510 and a 300 × 510 matrix to test!!!

Thus if \(d(L,S)\) is below \(μ-3σ\), we can safely assume that there
is statistically significant correlation between \(L\) and
\(S\). However, below, we will also give a sense of \emph{how} the
correlation is manifested by analysis of the two principal components
of correlation in the linguistic embeddings, \(U_0\) and \(U_1\). To
do so we will plot the projection of embeddings on along \(U_0\) and
\(U_1\), together with the corresponding singular values.

@@

%\begin{figure*}[t]
%\begin{tikzpicture}
  %\begin{groupplot}[
    %group style={group size=2 by 1},
    %yticklabels={,,}, xticklabels={,,},
    %legend style={anchor=north, legend columns=1, font=\sffamily\tiny},
  %]
    %\nextgroupplot
    %\commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
    %\nextgroupplot
    %\commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
    %\legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  %\end{groupplot}
%\end{tikzpicture}
%\caption{Plot of embeddings along the 2 principal dimensions of correlation between .... TODO of an LSTM (left, $c=1$) and transformer (right, $c=3$) model.
%Color correspond the 5-means clustering of .... TODO}
%\label{fig:comm-pca}
%\end{figure*}
